---
title: "Mini Data-Analysis Deliverable 3"
output: github_document
author: "Yue Shen"
date: "2021/10/26"
---


# Welcome to Yue Shen's last milestone in the mini data analysis project!



# Setup

Begin by loading your data and the tidyverse package below: 

```{r, message = FALSE}
suppressPackageStartupMessages(library(datateachr)) # <- might contain the data you picked!
suppressPackageStartupMessages(library(tidyverse))
```

From Milestone 2, you chose two research questions. What were they? Put them here.

<!-------------------------- Start your work below ---------------------------->
1. *Are trees with a specific same genus have longer diameter?*

2. *What's the trend of count of trees planted for different neighbourhood during the past year?*
<!----------------------------------------------------------------------------->


# Exercise 1: Special Data Types (10)

For this exercise, you'll be choosing two of the three tasks below -- both tasks that you choose are worth 5 points each. 

But first, tasks 1 and 2 below ask you to modify a plot you made in a previous milestone. The plot you choose should involve plotting across at least three groups (whether by facetting, or using an aesthetic like colour). Place this plot below (you're allowed to modify the plot if you'd like). If you don't have such a plot, you'll need to make one. Place the code for your plot below.

<!-------------------------- Start your work below ---------------------------->

In previous Research Question 4, I generated a summarized variable *date_range* from the variable ```date_planted```.

```{r}
RQ4_Summarize  <- vancouver_trees %>% mutate(date_range = case_when(
                               date_planted <= "1991-11-1" ~ "More than 10", 
                               date_planted <= "1996-11-1" ~ "Between 5 and 10",
                               date_planted <= "2021-11-1" ~ "Less than 5")) %>%
                 select(tree_id, date_planted, date_range, everything()) %>% drop_na()


RQ4_Summarize2 <- RQ4_Summarize %>% group_by(neighbourhood_name, date_range) %>% 
             summarise(tree_count = n(), .groups = 'drop') %>%
             arrange(neighbourhood_name, tree_count)

RQ4_Summarize2
```

Then I created a graph out of summarized variables *date_range* and trees' *count* in each neighbourhood. Also, I used facetting to distinguish the plot by neighbourhood. 

```{r}
RQ4_Graph <- RQ4_Summarize2 %>%
          ggplot(aes(date_range, tree_count)) +
          facet_wrap(~ neighbourhood_name) +
          geom_col() +
          xlab("Date Range when Planting Trees") +
          ylab("Number of Trees")
print(RQ4_Graph)
```
<!----------------------------------------------------------------------------->

Now, choose two of the following tasks.

1. Produce a new plot that reorders a factor in your original plot, using the `forcats` package (3 points). Then, in a sentence or two, briefly explain why you chose this ordering (1 point here for demonstrating understanding of the reordering, and 1 point for demonstrating some justification for the reordering, which could be subtle or speculative.)

2. Produce a new plot that groups some factor levels together into an "other" category (or something similar), using the `forcats` package (3 points). Then, in a sentence or two, briefly explain why you chose this grouping (1 point here for demonstrating understanding of the grouping, and 1 point for demonstrating some justification for the grouping, which could be subtle or speculative.)

3. If your data has some sort of time-based column like a date (but something more granular than just a year):
    1. Make a new column that uses a function from the `lubridate` or `tsibble` package to modify your original time-based column. (3 points)
        - Note that you might first have to _make_ a time-based column using a function like `ymd()`, but this doesn't count.
        - Examples of something you might do here: extract the day of the year from a date, or extract the weekday, or let 24 hours elapse on your dates.
    2. Then, in a sentence or two, explain how your new column might be useful in exploring a research question. (1 point for demonstrating understanding of the function you used, and 1 point for your justification, which could be subtle or speculative).
        - For example, you could say something like "Investigating the day of the week might be insightful because penguins don't work on weekends, and so may respond differently".
    


<!-------------------------- Start your work below ---------------------------->

**Task Number**: 1

- Firstly, I used function ```fct_reorder``` to reorder the factor ```neighbourhood_name``` in increasing order of their total count of trees, which are previously listed in their names' alphabetical order.  
As shown in the *Task1_Graph* below, the facets of all neighbourhoods have been ordered by total trees' count. The last sub-graph shows the neigbourhood which has the most trees. 

- Secondly, I used function ```fct_recode``` to change the factor levels of ```date_range``` since they're too wide for plotting. Now the x-axis in the *Task1_Graph* can show three levels ("yrs>10", "5-10", "yrs<5") clearly without overlapping.

- Thirdly, I also used function ```fct_relevel``` to rearrange the order of ```date_range```, because in the previous graph, the level of "Less than 5" is in the middle, which doesn't accord with the correct time order.  
Now, the three levels are rearranged in the correct order of "planted more than 10 years ago", "planted between 5 and 10 years" and "planted in recent 5 years".

```{r}
Task1 <- RQ4_Summarize2 %>%
           mutate(neighbourhood_name = fct_reorder(neighbourhood_name, tree_count, sum)) %>%
  
           mutate(date_range = fct_recode(date_range, "yrs>10" = "More than 10",
                                                      "5-10" = "Between 5 and 10",
                                                      "yrs<5" = "Less than 5")) %>%

           mutate(date_range = fct_relevel(date_range, "yrs>10", "5-10", "yrs<5"))
```



```{r}
Task1_Graph <- Task1 %>%
          ggplot(aes(date_range, tree_count)) +
          facet_wrap(~ neighbourhood_name) +
          geom_col() +
          xlab("Date Range when Planting Trees") +
          ylab("Number of Trees")
print(Task1_Graph)
```

<!----------------------------------------------------------------------------->




<!-------------------------- Start your work below ---------------------------->

**Task Number**: 3

```{r}
Task3  <- vancouver_trees %>%
          filter(!is.na(date_planted)) %>%
          mutate(year_planted = lubridate::year(date_planted)) %>%
          select(tree_id, year_planted, date_planted, everything()) 
Task3

```

I extracted the year from ```date_planted``` to create a new column named ```year_planted``` using the function in the ```lubridate``` package.

To explore my Research Question 4 "What's the trend of count of trees planted for different neighbourhood during the past year?", by extracting the year, it's more convenient for me to calculate how many trees are planted for each year, so that I could create the line graphs to explore the trend.



<!----------------------------------------------------------------------------->

    

# Exercise 2: Modelling

## 2.0 (no points)

Pick a research question, and pick a variable of interest (we'll call it "Y") that's relevant to the research question. Indicate these.

<!-------------------------- Start your work below ---------------------------->

**Research Question**: What's the trend of count of trees planted for different neighbourhood during the past year?

**Variable of interest**: ```tree_count``` as "Y"， ```year_planted``` as "X".

From the new data frame *Task3* which contains the extracted year, I created a new date frame *Exercise2_DF1* by grouping ```neighbourhood_name``` and ```year_planted```, to calculate the how many trees are planted in each year for different neighbourhoods.

```{r}
Exercise2_DF1 <- Task3 %>% group_by(neighbourhood_name, year_planted) %>% 
             summarise(tree_count = n(), .groups = 'drop')
Exercise2_DF1
```

<!----------------------------------------------------------------------------->

## 2.1 (5 points)

Fit a model or run a hypothesis test that provides insight on this variable with respect to the research question. Store the model object as a variable, and print its output to screen. We'll omit having to justify your choice, because we don't expect you to know about model specifics in STAT 545. 

- **Note**: It's OK if you don't know how these models/tests work. Here are some examples of things you can do here, but the sky's the limit. 
    - You could fit a model that makes predictions on Y using another variable, by using the `lm()` function.
    - You could test whether the mean of Y equals 0 using `t.test()`, or maybe the mean across two groups are different using `t.test()`, or maybe the mean across multiple groups are different using `anova()` (you may have to pivot your data for the latter two).
    - You could use `lm()` to test for significance of regression.

<!-------------------------- Start your work below ---------------------------->

I created a subset of the dataset *Exercise2_DF1* above containing only the neighbourhood of *DOWNTOWN* by filtering the ```neighbourhood_name```. 

```{r}
Exercise2_DF2 <- Exercise2_DF1 %>% filter(neighbourhood_name == "DOWNTOWN")
Exercise2_DF2
```

I used the ```lm()``` function to create the linear model of ```tree_count``` ("Y") and ```year_planted``` ("X").

```{r}
Exercise2.1 <- lm(tree_count ~ year_planted, Exercise2_DF2)
print(Exercise2.1)
```


<!----------------------------------------------------------------------------->

## 2.2 (5 points)

Produce something relevant from your fitted model: either predictions on Y, or a single value like a regression coefficient or a p-value. 

- Be sure to indicate in writing what you chose to produce. 
- Your code should either output a tibble (in which case you should indicate the column that contains the thing you're looking for), or the thing you're looking for itself.
- Obtain your results using the `broom` package if possible. If your model is not compatible with the broom function you're needing, then you can obtain your results by some other means, but first indicate which broom function is not compatible. 
  
<!-------------------------- Start your work below ---------------------------->

Firstly, I produced a *p-value* using the ```tidy()``` function in the ```broom``` package.

```{r}
Exercise2.2.1 <- broom::tidy(Exercise2.1)
print(Exercise2.2.1)
```

Then I predicted the ```tree_count``` of other neighbourhoods (except "DOWNTOWN") using the ``augment()``` function, as shown in the new generated column *.fitted*.

```{r}
Exercise2.2.2 <- Exercise2_DF1 %>%
          filter(neighbourhood_name != "DOWNTOWN") %>%
          select(!tree_count) %>%
          broom::augment(Exercise2.1, newdata = .)
print(Exercise2.2.2)
```


<!----------------------------------------------------------------------------->



# Exercise 3: Reading and writing data

Get set up for this exercise by making a folder called `output` in the top level of your project folder / repository. You'll be saving things there. 

## 3.1 (5 points)

Take a summary table that you made from Milestone 2 (Exercise 1.2), and write it as a csv file in your `output` folder. Use the `here::here()` function.

- **Robustness criteria**: You should be able to move your Mini Project repository / project folder to some other location on your computer, or move this very Rmd file to another location within your project repository / folder, and your code should still work.
- **Reproducibility criteria**: You should be able to delete the csv file, and remake it simply by knitting this Rmd file. 
  
<!-------------------------- Start your work below ---------------------------->

I chose the summary table I made in the Research Question 4 from Milestone 2 (Exercise 1.2), and the table name is *RQ4_Summarize*.

```{r}
RQ4_Summarize  <- vancouver_trees %>% mutate(date_range = case_when(
                               date_planted <= "1991-11-1" ~ "More than 10", 
                               date_planted <= "1996-11-1" ~ "Between 5 and 10",
                               date_planted <= "2021-11-1" ~ "Less than 5")) %>%
                 select(tree_id, date_planted, date_range, everything()) %>% drop_na()
RQ4_Summarize
```


```{r}
write_csv(RQ4_Summarize, here::here("output", "RQ4_Summarize.csv"))
```

<!----------------------------------------------------------------------------->




## 3.2 (5 points)

Write your model object from Exercise 2 to an R binary file (an RDS), and load it again. Be sure to save the binary file in your `output` folder. Use the functions `saveRDS()` and `readRDS()`. 

- The same robustness and reproducibility criteria as in 3.1 apply here.

<!-------------------------- Start your work below ---------------------------->

I chose the model I created in *Exercise2.1* and save it to an ```.rds``` file using the function ```saveRDS()```

After reloading it using the function ```readRDS()```, I can still get the original model object.

```{r}
saveRDS(Exercise2.1, here::here("output", "Exercise2.1.rds"))

Exercise2.1_reload <- readRDS(here::here("output", "Exercise2.1.rds"))
print(Exercise2.1_reload)
```

<!----------------------------------------------------------------------------->
