---
title: "Mini Data-Analysis Deliverable 2"
output: github_document
author: "Yue Shen"
date: "2021/10/17"
---


# Welcome back to Yue Shen's mini data analysis project!

Begin by loading your data and the tidyverse package below: 

```{r, message = FALSE}
suppressPackageStartupMessages(library(datateachr))
suppressPackageStartupMessages(library(tidyverse))
```

# Learning Objectives

By the end of this milestone, you should:

+ Become familiar with manipulating and summarizing your data in tibbles using `dplyr`, with a research question in mind
+ Understand what *tidy* data is, and how to create it. In milestone 3, we will explore when this might be useful
+ Generate a reproducible and clear report using R Markdown
+ Gain a greater understanding of how to use R to answer research questions about your data



# Task 1: Process and summarize your data (15 points)


### 1.1 (2.5 points) 

First, write out the 4 research questions you defined in milestone 1 were. This will guide your work through milestone 2:

<!-------------------------- Start your work below ---------------------------->

1. Does a particular street have more trees than others?

2. Are trees with a specific same species planted in a range of latitude?

3. Are trees with a specific same genus have longer diameter?

4. For different neighbourhood, what range of years have the highest frequency of planting trees?

<!----------------------------------------------------------------------------->

### 1.2 (10 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question. 

Ensure that the output of each operation is printed! 

**Summarizing:**

1. Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2. Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
3. Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4. Based on two categorical variables, calculate two summary statistics of your choosing.

**Graphing:**

5. Create a graph out of summarized variables that has at least two geom layers.
6. Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read. 
7. Make a graph where it makes sense to customize the alpha transparency.
8. Create 3 histograms out of summarized variables, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

<!------------------------- Start your work below ----------------------------->

#### 1. Does a particular street have more trees than others?

**Summarizing:**

Choose 2: I computed the number of observations for categorical variables ```std_street``` and arrange them in a descending order.

```{r}
RQ1_Summarize  <- vancouver_trees %>% group_by(std_street) %>% summarize(tree_count = n()) %>% arrange(desc(tree_count))
RQ1_Summarize
```

**Graphing:**

Choose 8: I created 3 histograms out of the summarized variables which shows number of tree in different streets. For 3 histograms, I set the bar size 5, 30 and 100 respectively.
```{r}
RQ1_Graph <- ggplot(RQ1_Summarize, aes(tree_count)) +
   geom_histogram(bins = 5) +
   xlab("Number of Trees per Street") +
   ylab("Count of Streets")
print(RQ1_Graph)
```
```{r}
RQ1_Graph <- ggplot(RQ1_Summarize, aes(tree_count)) +
   geom_histogram(bins = 30) +
   xlab("Number of Trees per Street") +
   ylab("Count of Streets")
print(RQ1_Graph)
```



```{r}
RQ1_Graph <- ggplot(RQ1_Summarize, aes(tree_count)) +
   geom_histogram(bins = 100) +
   xlab("Number of Trees per Street") +
   ylab("Count of Streets")
print(RQ1_Graph)
```

I personally think the third histogram with bar size 100 is the "best" one, because we could clearly see that there are only 2 streets with more than 1500 trees, and few streets have more than 1000 trees. The majority of the streets own trees with less than the number of 500.

However, since there exists a large amount of different street, the bars with small size will be hard to distinguished. 




#### 2. Are trees with a specific same species planted in a range of latitude?

**Summarizing:**

Choose 1: I computed the *range*, *mean*, *median*, and *count* of trees across the groups of ```species_name```, and arrange them in the ascending order of range, mean and count.

```{r}
RQ2_Summarize  <- vancouver_trees %>% group_by(species_name) %>% drop_na() %>% 
        summarise(min_latitude = min(latitude), max_latitude = max(latitude), range = max_latitude - min_latitude,
                  mean = mean(latitude, na.rm = TRUE), 
                  median = median(latitude, na.rm = TRUE), count  = n()) %>% 
                  arrange(range, mean, count) 
RQ2_Summarize
```


**Graphing:**

Choose 5 and 7: I created a graph out of summarized variables *mean* of latitude and trees' *count* for each species. Also, it's a plot with multiple points, some of which have possible overlapping positions, hence, I customized the alpha transparency to 0.3, and set point color to red.

```{r}
RQ2_Graph <- ggplot(RQ2_Summarize, aes(mean, count)) +
   geom_point(color = "red", alpha = 0.3) +
   geom_line() +
   xlab("Mean of Tree's Diameter for Different Species") +
   ylab("Number of Trees")
print(RQ2_Graph)
```



#### 3. Are trees with a specific same genus have longer diameter?

**Summarizing:**

Choose 1: I computed the *range*, *mean*, and *count* of trees and *count* of trees whose mean diameter is larger than 5, across the groups of ```genus_name```, and arrange them in the ascending order of mean and count.

```{r}
RQ3_Summarize  <- vancouver_trees %>% group_by(genus_name) %>% drop_na() %>% 
        summarise(min_diameter = min(diameter), max_diameter = max(diameter), range = max_diameter - min_diameter,
                  mean_diameter = mean(diameter, na.rm = TRUE), 
                  tree_count  = n(), tree_count_5 = sum(diameter >= 5)) %>% 
                  arrange(mean_diameter, tree_count)
RQ3_Summarize
```


**Graphing:**

Choose 6: I created a graph with mean diameter as x-axis and count of trees as y-axis. I made y-axis logarithmic, and formatted its label as "Number of  trees". I also colored those points by genus name. Making y-axis logarithmic helps those points become more scattered and distinguishable.

```{r}
RQ3_Graph <-  RQ3_Summarize %>% 
          ggplot(aes(mean_diameter, tree_count)) +
          geom_point(size = 2.5, aes(color = genus_name)) +
          xlab("Mean of Tree's Diameter") +
          scale_y_log10("Number of Trees")
print(RQ3_Graph)
```

#### 4. For different neighbourhood, what range of years have the highest frequency of planting trees?

**Summarizing:**

Choose 3: I created a categorical variable```date_range``` with 3 groups from existing numerical variable ```date_planted```. 

- If the tree is planted before 1991-11-1. then it should be planted for more than 10 years;

- If the tree is planted before 1996-11-1. then it should be planted for more than 5 years but less than 10 years;

- If the tree is planted before 2021-11-1. then it should be planted for less than 5 years;

```{r}
RQ4_Summarize  <- vancouver_trees %>% mutate(date_range = case_when(
                               date_planted <= "1991-11-1" ~ "More than 10", 
                               date_planted <= "1996-11-1" ~ "Between 5 and 10",
                               date_planted <= "2021-11-1" ~ "Less than 5")) %>%
                 select(tree_id, date_planted, date_range, everything()) %>% drop_na()
RQ4_Summarize
```

After adding the new column ```date_range```, I calculated the tree count based on it and another categorical variables ```neighbourhood_name```. 

```{r}
RQ4_Summarize2 <- RQ4_Summarize %>% group_by(neighbourhood_name, date_range) %>% 
             summarise(tree_count = n(), .groups = 'drop') %>%
             arrange(neighbourhood_name, tree_count)
RQ4_Summarize2
```

**Graphing:**

Choose 5: I created a graph out of summarized variables *date_range* and trees' *count* in each neighbourhood. Also, I used facetting to distinguish the plot by neighbourhood. 

```{r}
RQ4_Graph <- RQ4_Summarize2 %>%
          ggplot(aes(date_range, tree_count)) +
          facet_wrap(~ neighbourhood_name) +
          geom_col() +
          xlab("Date Range when Planting Trees") +
          ylab("Number of Trees")
print(RQ4_Graph)
```


<!----------------------------------------------------------------------------->


### 1.3 (2.5 points)


Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?

<!------------------------- Write your answer here ---------------------------->


**RQ1:** 
I think I mostly figure out what I want to know. I found the street with most trees! Even though there are too many streets to draw a more clear graph, it still tells us that there are only 2 streets with more than 1500 trees, and few streets have more than 1000 trees. The majority of the streets own trees with less than the number of 500.

**RQ2:** 
I don't think the relationship between latitude and tree's species is clear, since the dataset contains trees only in Vancouver, the range of latitude is not big. I can only see that some of the species in Vancouver is very rare.

**RQ3:** 
The results are interesting. We can clearly see that the trees with some of genus have longer diameter, like "ROBINIA" and "PLATANUS".

**RQ4:** 
From the research above, We can see that most of the neighbourhoods have the majority trees planted during the recent 5 years. 

However, I couldn't distinguish the frequency of planting trees in different neighbourhood clearly, because I only splitted the ```date_planted``` into 3 groups.

For future investigation, maybe the research question can be refined as "What's the trend of count of trees planted for different neighborhood during the past year?"


<!----------------------------------------------------------------------------->







# Task 2: Tidy your data (12.5 points)

In this task, we will do several exercises to reshape our data. The goal here is to understand how to do this reshaping with the `tidyr` package.

A reminder of the definition of *tidy* data:

- Each row is an **observation**
- Each column is a **variable**
- Each cell is a **value**

*Tidy'ing* data is sometimes necessary because it can simplify computation. Other times it can be nice to organize data so that it can be easier to understand when read manually. 

### 2.1 (2.5 points)

Based on the definition above, can you identify if your data is tidy or untidy? Go through all your columns, or if you have >8 variables, just pick 8, and explain whether the data is untidy or tidy.

<!--------------------------- Start your work below --------------------------->
```{r}
my_vancouver_trees <- vancouver_trees %>% select(tree_id, std_street, genus_name, species_name, neighbourhood_name, diameter, date_planted, latitude)
my_vancouver_trees
```

I picked 8 variables that I've used. The data is tidy because it satisfies the three conditions above:

- Each row is an observation of one tree.

- Each column is a variable which describes one characteristic of tree. Different trees have different id, locations, genus, diameter......

- Each cell shows an exact value of that tree.


<!----------------------------------------------------------------------------->

### 2.2 (5 points)

Now, if your data is tidy, untidy it! Then, tidy it back to it's original state.

If your data is untidy, then tidy it! Then, untidy it back to it's original state.

Be sure to explain your reasoning for this task. Show us the "before" and "after".

<!--------------------------- Start your work below --------------------------->

### Before - the data is tidy
```{r}
my_vancouver_trees
```

### Untidy my_vancouver_trees

```{r}
untidy_vancouver_trees <- my_vancouver_trees %>%
    unite(col = "name", c(genus_name, species_name), sep = "_" )
untidy_vancouver_trees
```
I united ```genus_name``` and ```species_name``` into a new variable ```name``` using ```tidyr::unite()``` and joined then with "_", which made the data untidy because each cell contained more than one value.


### After - tidy back
```{r}
tidy_vancouver_trees <- untidy_vancouver_trees %>%
    separate(name, into = c("genus_name", "species_name"), sep="_")
tidy_vancouver_trees
```
I used ```tidyr::separate()``` to split the ```name``` in ```untidy_vancouver_trees``` back into the original two columns: ```genus_name``` and ```species_name```, making the data tidy again!

<!----------------------------------------------------------------------------->

### 2.3 (5 points)

Now, you should be more familiar with your data, and also have made progress in answering your research questions. Based on your interest, and your analyses, pick 2 of the 4 research questions to continue your analysis in milestone 3, and explain your decision. 

Try to choose a version of your data that you think will be appropriate to answer these 2 questions in milestone 3. Use between 4 and 8 functions that we've covered so far (i.e. by filtering, cleaning, tidy'ing, dropping irrelevant columns, etc.). 

<!--------------------------- Start your work below --------------------------->

I want to choose question 3 and question 4 to continue my analysis.

- Question 3: Are trees with a specific same genus have longer diameter?

- Question 4: For different neighbourhood, what range of years have the highest frequency of planting trees? (Refined as: What's the trend of count of trees planted for different neighborhood during the past year?")


Explanation: For questin 3, It's more valuable to dig into the tree's genus and their diameters, which may produce many interesting academic findings.

Also, for question 4, we can add a new column into the dataframe, and through analyzing this question, we can have a better understanding of each neighborhood and its types of trees, which many people especially residents may be interested in.


I used function ```drop_na()``` to drop irrelevant columns, function ```arrange``` to sort the tree by its date planted, function ```mutate``` to add a new column, function ```rename``` to rename the title of columns, and function ```select``` to select the columns that I need.

```{r}
new_vancouver_trees  <- vancouver_trees %>% 
                      drop_na() %>% 
                      arrange(date_planted) %>% 
                      mutate(date_range = case_when(
                               date_planted <= "1991-11-1" ~ "More than 10", 
                               date_planted <= "1996-11-1" ~ "Between 5 and 10",
                               date_planted <= "2021-11-1" ~ "Less than 5")) %>%
                      rename(genus = genus_name, species = species_name, 
                            neighbourhood = neighbourhood_name, street_side = street_side_name) %>%
                      select(tree_id, date_planted, date_range, std_street, genus, species, neighbourhood, street_side, height_range_id, diameter, longitude, latitude)
new_vancouver_trees
```

I used function ```saveRDS()``` to save the new dataframe as an ```.RDS``` file.


```{r}
saveRDS(new_vancouver_trees,file="new_vancouver_trees")
```

<!----------------------------------------------------------------------------->

*When you are done, knit an `md` file. This is what we will mark! Make sure to open it and check that everything has knitted correctly before submitting your tagged release.*

### Attribution

Thanks to Victor Yuan for mostly putting this together. 